---
title: "Velo.com A/B Test Revisited"
author: "Natalie Benoy"
date: "10/23/21"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

<!-- Note:   -->

<!-- These instructions are commented out and will not display when you knit your RMarkdown document. -->

<!-- - Change the information in the yaml header above:  title, author, data. -->
<!-- - Make sure output is html_document. -->
<!-- - Once you are finished coding, **run each chunk individually to make sure there are no errors**.  (If necessary fix your code.) Once your code is error-free, click "knit" on the menu above. Your document should compile to HTML, provided that you have output set to "html_document." -->
<!-- - In the code chunk above ("setup") echo is set to TRUE.  This means that the code in your chunks will be displayed, along with the results, in your compiled document. -->

## Load packages and create empty data frame

```{r}
library(tidyverse)
v <- read_csv("velo.csv")

```

Note that `lm()` will automatically assign alphabetic factor levels to character predictors.  To make the regression coefficient for checkout_system easier to interpret in this tutorial we will manually change the factor levels.  This will ensure that the coefficient is, appropriately, positive.

```{r}

v <- v %>% 
  mutate(checkout_system = factor(checkout_system, 
                                  levels = c("old", "new"))) # or

# v$checkout_system <- factor(v$checkout_system, levels = c("old", "new"))

```

## Questions

The point of this case is to revisit velo.com  with a different methodology.  For reference, in the previous module you we obtained the following results from a t-test comparing the old and new systems:

```{r}
# T-test comparing old and new systems
t.test(filter(v, checkout_system == "new")$spent, 
       filter(v, checkout_system == "old")$spent)

# T-test comparing old and new systems for mobile users
t.test(filter(v, checkout_system == "new" & device == "mobile")$spent, 
       filter(v, checkout_system == "old" & device == "mobile")$spent)

```

### Q1

Fit a simple linear regression with `spent` as the target variable and `checkout_system` as the predictor. Based on the model summary, answer the following questions:

1. What is average spending for customers using 1) the new system and 2) the old system? This information can be extracted directly from the linear model output.

2. What is the difference in customer spending between the new and old systems, and is the difference statistically significant at the p < .05 level?

3. Compare these results to those you obtained using a t-test in the last module (see the above output).

```{r}
# fit the model & save as object
(model1 <- lm(spent ~ checkout_system, data = v))

# double check average spent by checkout system
v %>%
  group_by(checkout_system) %>%
  summarize(average = mean(spent))

# let's get a summary of our model
summary(model1)

# calculate 95% CI for coefficient - does it include 0?
14.855 - 1.96 * 11.133
14.855 + 1.96 * 11.133

```

> Answer:  

> 1. Regression equation is: spent = 1956.39 + 14.85 * checkout_system. When checkout_system = old, average amount spent is 1956.39. When checkout_system = new, average amount spend is 1971.24. Used summary table to double-check these values.

> 2. The difference between the old and new checkout systems is, on average, $14.85 in spending. The p-value returned for a two-sided, two-tailed t-test is .18, which means this result is not statistically significant at the p < .05 level. A 95% CI for the coefficient ranges from -6.97 to 36.68, which includes 0 and is therefore not statistically significant.

> 3. Both the p-value and 95% CI returned by the linear model are virtually identical to those given by the t-test function above. More importantly, the conclusion (that there is not a statistically significant difference in amount spent by users of the new system compared to the old system) is the same from these results as it is from the results of the t-test above.

### Q2

Fit a simple linear regression with `spent` as the target variable and `checkout_system` as the predictor, but include only customers who were using a mobile device. (Later you will learn how to fit this sort of regression using all the data, but for now subset the data to include only mobile users.) Answer these questions based on the model summary for this regression:

1. What is the difference in mobile customer spending between the new and old systems?

2. Is the difference statistically significant at the p < .05 level?

3. Compare these results to those you obtained using a t-test in the last module (see the above output).


```{r}
# fit a model for spent ~ checkout_system for mobile users only, save as object
model2 <- lm(spent ~ checkout_system, data = filter(v, device == "mobile"))

summary(model2)

# use summary table to double-check
v %>%
  group_by(checkout_system, device) %>%
  summarize(average = mean(spent))

# calculate 95% CI for coefficient
35.19 - 1.96 * 16.16
35.19 + 1.96 * 16.16
```

> Answer:  

> 1. Regression equation of spent ~ checkout system for mobile users is: spent = 1931.69 + 35.19 * checkout_system. When checkout_system = old, average amount spent is 1931.69. When checkout_system = new, average amount spent is 1966.88. Used summary table to double-check these values.

> 2. The difference between the old and new checkout systems is, on average, $35.19 in spending. The p-value returned for a two-sided, two-tailed t-test is .0295, which means this result is statistically significant at the p < .05 level. A 95% CI for the coefficient ranges from 3.52 to 66.86, which does not include 0 and is therefore considered statistically significant.

> 3. Both the p-value and 95% CI returned by the linear model are virtually identical to those given by the t-test function above. More importantly, the conclusion (that there is a statistically significant difference in amount spent by mobile users of the new system compared to the old system) is the same from these results as it is from the results of the t-test above.

###  Q3

Using the summary of the regression model object from Q2, calculate a 95% confidence interval (CI) for the `checkout_system` coefficient using 1.96 as the critical value. Interpret the CI: What does it mean exactly?

```{r}
# oops already did this, going to reprint it here:

# calculate 95% CI for coefficient
35.19 - 1.96 * 16.16
35.19 + 1.96 * 16.16
```

> Answer: A 95% CI for the checkout_system coefficient ranges from 3.52 to 66.86, which does not include 0 and is therefore considered statistically significant. This CI represents the range of values the checkout_system coefficient would take 95% of the time under repeated sampling. The fact that the CI does not include 0 leads us to believe there is a non-zero difference in mean amount spent by checkout system.

### Q4

Based on the model output (and equation) in Q2, along with the 95% CI you calculated in Q3, develop best and worst case revenue projections for the new checkout system.  Assume that you will see 10,000 mobile users in a given month, and use revenue under the old system as the baseline. What range of increased revenue might the company expect and, based on this, does the coefficient estimate for `checkout_system`  have *practical* significance in your view?

A 95% CI for a coefficient represents the most likely range of values under repeated sampling.  Working with a range is advantageous.  Velo.com executives might well be interested in the low end of the range in order to assess the financial risks associated with implementing the new system.  Using just the coefficient tends to obscure this range and might make future expectations of revenue overly rosy. 

The CI, like the coefficient, is expressed in terms of dollars per customer.  To put the CI in terms of dollars per 10,000 customers you need to multiply the lower and upper bounds appropriately.  You do not need to use the regression equation.  Simply use the interval you calculated in Q3.

```{r}
# reprinting 95% CI for checkout_system coefficient
35.19 - 1.96 * 16.16
35.19 + 1.96 * 16.16

# calculate baseline average revenue under old system for 10,000 users
1931.68 * 10000

# calculate lower bound & upper bound average revenue for new system for 10,000 users
(1931.68 + 35.19 - 1.96 * 16.16) * 10000
(1931.68 + 35.19 + 1.96 * 16.16) * 10000

# or, remove baseline and focus only on value-add for new system
(35.19 - 1.96 * 16.16) * 10000
(35.19 + 1.96 * 16.16) * 10000
```

> Answer: Compared to our baseline average revenue of $19,316,800 for 10,000 mobile users, we can expect average revenue under the new system to range between $19,351,964 on the low end and $19,985,436 on the high end, 95% of the time. Another way of saying this is that the new system will add between $35,164 and $668,636 in revenue per 10,000 users, on average.

### Q5

What course of action should Sarah recommend to the management at velo.com based on this additional analysis? Please incorporate your analytic results from above in fashioning an answer. In particular, cite the results from the new analysis, the simple linear regression, and discuss your interpretation of the range you developed in Question 4.

> Regarding the results of the recent A/B test comparing average amount spent by users of the new vs. old checkout systems, we recommend velo.com adopt the new checkout system for mobile users only. After using simple linear regression to determine the relationship between checkout system and amount spent, we found an effect size of 35.19 under the new system -- that is, mobile users of the new checkout system spent on average $35.19 more per transaction than mobile users of the old checkout system. When we compared the results of this simple linear regression against a t-test comparing average spending of mobile users of the old vs. new checkout systems, we found both analyses supported the same conclusion. Mobile users of the new checkout system spent more than users of the old system; this result was statistically significant at the p < .05 level, and is therefore unlikely to have occurred by chance.

> Our analysis indicates that, per 10,000 monthly mobile users, we can expect to generate an additional $35,164 - $668,636 in revenue using the new system compared to the old, with an average expected value-add of $351,900 (put another way, we expect the new system to generate between $19,351,964 and $19,985,436 in total monthly mobile revenues, compared to our current revenue baseline of $19,316,800 for 10,000 mobile users). Thus, even at the lower end of expected values, the new system is clearly superior in its ability to generate revenue from mobile transactions.

> In addition, there is some evidence the new system results in fewer cart abandonments on mobile — in order to compare spending, our dataset included only users who actually completed a transaction. Since users were randomly assigned to either the new or old system, we should expect to see roughly an equivalent number of users in each condition in this dataset. However, while we observed 11495 completed transactions for the new system on mobile, we observed only 9663 transactions for the old system during this period. This difference is not observed among desktop computer users. We recommend further analysis of mobile transaction completion rates under the new system going forward to determine if there is a further economic benefit to the new system that has not been captured in this analysis.

### Challenge

Read through the tutorial on the bootstrap (included in the lightbulb package).  Use the technique to estimate a confidence interval for `checkout_system` regression coefficient.  Does it agree with the confidence interval computed using the SEs for the coefficient from the regression model output?  Please show your code for your calculations.

```{r}
# Initialize empty vector
boot_distribution <- NULL

# Specify 1000 simulations
sims <- 1000

# Subset data by old vs new
old <- filter(v, device == "mobile", checkout_system == "old")
new <- filter(v, device == "mobile", checkout_system == "new")

# Calculate the difference in median amount spent for each iteration
set.seed(123)

for(i in 1:sims){
  
  boot_old <- sample(old$spent, replace = T)
  boot_new <- sample(new$spent, replace = T)
  
  # These are both temporary bootstrap samples that will be overwritten/redefined
  # for every loop.
  
  boot_distribution[i] <- median(boot_new) - median(boot_old)
  
  # We index the vector, boot_distribution, with [i] in order to accumulate
  # a new simulated difference for each loop iteration.
  
}

# Inspect the bootstrap distribution of difference in amount spent
head(boot_distribution, 20)

# Calculate lower and upper bounds for 95% CI
mean(boot_distribution) - 1.96 * sd(boot_distribution)
mean(boot_distribution) + 1.96 * sd(boot_distribution)
```
> The confidence interval has been shifted slightly to the right. Why? ...good question.